### 2018-06-05 Physics Notes/Paper 1/Paper 1.tex

@@

The lower the temperature, the less the vibration of the atoms, and resistance reduces

@@

>>>>

\pgfkeys{/pgfplots/Axis Style 2/.style={
    width=13.5cm, height=5cm,
    axis x line=center, 
    axis y line=middle, 
    samples=100,
    ymin=-1.5, ymax=1.5,
    xmin=0, xmax=14.0,
    domain=-4*pi:4*pi
}}

@@

The lower the temperature, the less the vibration of the atoms, and resistance reduces\\

@@

\section{Periodic motion}
\subsection{Circular motion}
As circular motion has a constant speed but a changing direction, it is implied that there is a force perpendicular to the direction of motion, and so an acceleration\\
$v$ (Velocity)- Displacement per unit time\\
$\omega$ (Angular velocity)- Angle per unit time\\
\subsection{Simple harmonic motion}
    Axis Style 2,
    xtick={
        3.14159,6.28319,9.42478,12.56637
    },
    xticklabels={
        $\dfrac{T}{2}$,T,$\dfrac{3T}{2}$,2T
    },
    ytick={-1,1},
    yticklabels={-A,A},
    y label style={at={(axis description cs:-0.1,.5)},rotate=90,anchor=south},
    ylabel = {Displacement, x},
    xlabel= {Time, t}
\addplot [mark=none, ultra thick, black] {cos(deg(x))};
$x=Acos(\omega t)$
    Axis Style 2,
    xtick={
        3.14159,6.28319,9.42478,12.56637
    },
    xticklabels={
        $\dfrac{T}{2}$,T,$\dfrac{3T}{2}$,2T
    },
    ytick={-1,1},
    yticklabels={$-\omega A$,$\omega A$},
    y label style={at={(axis description cs:-0.1,.5)},rotate=90,anchor=south},
    ylabel = {Velocity, v},
    xlabel= {Time, t}
\addplot [mark=none, ultra thick, black] {-sin(deg(x))};
$v=-\omega A\sin(\omega t)$
    Axis Style 2,
    xtick={
        3.14159,6.28319,9.42478,12.56637
    },
    xticklabels={
        $\dfrac{T}{2}$,T,$\dfrac{3T}{2}$,2T
    },
    ytick={-1,1},
    yticklabels={$-\omega^2 A$,$\omega^2 A$},
    y label style={at={(axis description cs:-0.1,.5)},rotate=90,anchor=south},
    ylabel = {Acceleration, a},
    xlabel= {Time, t}
\addplot [mark=none, ultra thick, black] {-cos(deg(x))};
$a=-\omega^2A\cos(\omega t)$\\
All these equations can be derived from the equation $x=A\cos(\omega t)$, by differentiating it with respect to time.\\
The maximum values of each of them can be found as the maximum value of cos or sin is 1
\subsection{Simple harmonic systems}
\subsubsection{Energy in SHM}
    Axis Style 2,
    xtick={
        3.14159,6.28319
    },
    xticklabels={
        $\dfrac{T}{2}$,T
    },
    ytick={1},
    yticklabels={$\frac{1}{2}m\omega^2A^2$},
    y label style={at={(axis description cs:-0,.25)},rotate=90,anchor=south},
    ylabel = {Energy/J},
    xlabel= {Time, t},
    x label style={at={(axis description cs:1.12,0)},anchor=east},
    xmax=6.28319,
    ymin=0
\addplot [mark=none, ultra thick, black,samples=200] {(cos(deg(x)))^2};
\addplot [mark=none, ultra thick, red,samples=200] {(sin(deg(x)))^2};
Red - Kinetic Energy\\
Black - Gravitational potential energy\\
\begin{axis}[ymin=0,ticks=none,xlabel={Displacement,x/m},ylabel={Energy/J}]
\addplot [mark=none, ultra thick, black,samples=200] {x^2};
\addplot [mark=none, ultra thick, red,samples=200] {-x^2+25};
\draw[ultra thin] (axis cs:0,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:0,\pgfkeysvalueof{/pgfplots/ymax});
\subsubsection{Period of a mass-spring system}
The period of a mass-spring system is $T=2\pi\sqrt{\dfrac{m}{k}}$ or $T^2=\dfrac{4\pi^2m}{k}$\\
\begin{tikzpicture}[scale=0.5]
\begin{axis}[ymin=-0.001,ticks=none,ylabel=$T^2/s^2$,xlabel=$m/kg$,axis lines=middle,
label style={font=\large}
\addplot[mark=none]{x};
The gradient is $\dfrac{4\pi^2}{k}$\\
The line passes through the origin\\
This is only true for small oscillations as large oscillations go past the limit of proportionality
\subsubsection{Energy in a vertically oscillating spring}
\begin{tabular}{|c|c|c|c|}
&KE&GPE&Elastic Energy\\
Top&0&$2mgA$&$\frac{1}{2}k(e-A)^2$\\
Middle&Max=$\frac{1}{2}m\omega^2A^2$&$mgA$&$\frac{1}{2}ke^2$\\
Bottom&0&0&$\frac{1}{2}k(e+A)^2$\\

\subsubsection{Damping}
Damping is the reduction in amplitude due to energy losses (e.g. overcoming friction).\\
\\
It will also cause the period to increase (it is essentially a driver).\\
\\
\textbf{Light(underdamping)} - Small amplitude change per oscillation\\
\textbf{Heavy(underdamping)} - Large amplitude change per oscillation\\
\\
\textbf{Critical damping} - Reaches equilibrium in the shortest possible time \textbf{without} oscillating\\
\\
\textbf{Over damping} - Reaches equilibrium after a long period of time
\paragraph{Damping graph}
\begin{tikzpicture}
\begin{axis}[axis lines=middle,samples=200,ymax=10,ticks=none,ylabel=Amplitude,ylabel style={rotate=90, anchor=south,at={(axis description cs:0,0.42)},},xlabel=Time,x label style={at={(axis description cs:1.1,0.42)},anchor=east},]
\addplot[blue,domain=0:21] {1/x};\addlegendentry{Critically Damped}
\addplot[red,domain=0.01:20.58] {78*sin(deg(x+7.7))/(x+7.7)};\addlegendentry{Under Damped}
\addplot[green,domain=0.01:21] {-3*ln(x+2)+12};\addlegendentry{Over damped}
\end{axis}
\end{tikzpicture}
\subsection{Forced vibrations and resonance}
\textbf{Free vibrations} - Vibrations where there is no damping and no periodic force acting on the system, amplitude of the oscillations is constant\\
\textbf{Forced vibrations} - Vibrations of a system subjected to an external periodic force
\subsubsection{The effects of damping on the sharpness of resonance}
\includegraphics[width=8cm]{damp_6.jpg}\\
The less damping there is, the sharper the resonance
\subsubsection{Standing waves}
Standing waves can be produced at the resonant (natural) frequencies of a material, causing point on it to alternate between minimum and maximum amplitude, causing solid materials such as concrete to break.
